---
title: "Final_Code"
author: "Yue Cao"
date: "12/4/2017"
output: html_document
---
##Load data
```{r}
library(dplyr)
library(psych)
library(lmtest)
setwd("~/Desktop/statistical consulting/assignment/final project")
parkin = read.csv("./data/Liana_PD\ project.csv")
# 2961
parkin = filter(parkin, RequiredFieldsSiteName != "University of Florida (Gainesville)") %>% filter(visit == 0) %>% filter(NeuroExamPrimaryDiagnos %in% c("Control", "Parkinson's disease"))
# 1335
colnames(parkin)
pa = parkin[,c(2, 9, 10, 13,14, 15, 16, 17, 21:29)] # include duration 
```

##Check data
```{r}
summary(pa$duration) #670 missing values
pa$duration[which(pa$NeuroExamPrimaryDiagnos == "Control")] <- 0
pa <- pa[which(pa$duration >= 0 | is.na(pa$duration)),] # remove negative duration obs

summary(pa$duration) # 167 missing values
sum(pa$NeuroExamPrimaryDiagnos == "Parkinson's disease") # 830 PD, 20% missing values 
hist(pa$duration[pa$NeuroExamPrimaryDiagnos == "Parkinson's disease"], breaks = 40) 

#check if duration is missing at random in the case group
casegroup <- pa[pa$NeuroExamPrimaryDiagnos== "Parkinson's disease",]
summary(casegroup$duration) # why duration < 0??
casegroup$duration[which(casegroup$duration<0)] <- 0
durationmissing <- ifelse(!is.na(casegroup$duration), 1, 0)
fit1 <- glm(durationmissing ~ casegroup$MOCA_Total, family = "binomial")
summary(fit1) # the missing of duration is related to the MOCA score


```

##Descriptive data analysis
```{r}
######### divide groups by MOca score, under complete cases ######
p = parkin[,c(2, 9, 10, 14, 15, 16, 17, 21:29)] # not include duration
p = p[complete.cases(p),] #1177 obs
p = filter(p, p$Hgtcm > 100)  #1171obs
# combine education categories
p$edu[p$edu=="1"] <- "2"
p$edu <- factor(p$edu)
# calculate BMI
p$BMI = p$Wgtkg/(p$Hgtcm/100)^2

p1 <- p[which(p$MOCA_Total <= 19),]
p2 <- p[which(p$MOCA_Total >= 20 & p$MOCA_Total < 26),]
p3 <- p[which(p$MOCA_Total >= 26),]

## check BMI
plot(p$BMI,p$MOCA_Total)
summary(p1$BMI)
summary(p2$BMI)
summary(p3$BMI)

##continuous normal: age, height, weight, sbp, dbp, depression score, motor score, moka
describe(p1$Age)
describe(p2$Age)
describe(p3$Age)
 

# categorical: diagnos, gender, race, edu, smoke, med, diabetes
#proportion
table(p1$NeuroExamPrimaryDiagnos)/nrow(p1)




# continuous normal: age, height, weight, sbp, dbp
# continous skew: depression score, motor score, moka
hist(p$Age)
summary(p$Age)
hist(p$HDRSTotScore)
hist(p$MotorExamination_PartIIIScore)
summary(p$Hgtcm)
p = filter(p, p$Hgtcm > 100)
# 1171
hist(p$Hgtcm)
hist(p$Wgtkg)
hist(p$avg_SBP)
hist(p$avg_DBP)
hist(p$MOCA_Total)

summary(p$duration)
# categorical: diagnos, gender, race, edu, smoke, med, diabetes


```

## Calculate Framingham Score
```{r}

X.raw = p[,c(4, 14, 2, 12, 7, 17, 15)]
Fram <- function(X.raw){
  Framingham <- rep(NA, nrow(X.raw))
  for(i in 1:nrow(X.raw)){
    X <- X.raw[i,]
    X$Age = log(X.raw[i,]$Age)
    X$avg_SBP = log(X.raw[i,]$avg_SBP)
    X$BMI = log(X.raw[i,]$BMI)
    
    beta.male = data.frame(Age = 3.11, avg_SBP.NoTRT = 1.85, avg_SBP.TRT = 1.93,
                           Smoke_Ever = 0.71, BMI = 0.79, diabetes = 0.53)
    beta.female = data.frame(Age = 2.72, avg_SBP.NoTRT = 2.81, avg_SBP.TRT = 2.88,
                             Smoke_Ever = 0.62, BMI = 0.51, diabetes = 0.78)
    
    if (X$Gender == 0) {
      Xbeta = sum(X[, c('Age', 'Smoke_Ever', 'BMI', 'diabetes')] *
                    beta.male[, c('Age', 'Smoke_Ever', 'BMI', 'diabetes')])
      if (X$HTN_meds > 0) {
        Xbeta = Xbeta + X$avg_SBP * beta.male$avg_SBP.TRT
      } else {
        Xbeta = Xbeta + X$avg_SBP * beta.male$avg_SBP.NoTRT
      }
      Framingham[i] = 1 - 0.88431 ^ exp(Xbeta - 23.9388)
    } else {
      Xbeta = sum(X[, c('Age', 'Smoke_Ever', 'BMI', 'diabetes')] *
                    beta.female[, c('Age', 'Smoke_Ever', 'BMI', 'diabetes')])
      if (X$HTN_meds > 0) {
        Xbeta = Xbeta + X$avg_SBP * beta.female$avg_SBP.TRT
      } else {
        Xbeta = Xbeta + X$avg_SBP * beta.female$avg_SBP.NoTRT
      }
      Framingham[i] = 1 - 0.94833 ^ exp(Xbeta - 26.0145)
    }
  }
  return(Framingham)
}

p$framing = Fram(X.raw)

#rescale framingham score
p$framing <- 10* p$framing



```

### Explore linearity between dependent variables and the independent variable
```{r}
# age, height, weight, sbp, dbp
# depression score, motor score, moka
#aim 1
linearity <-  function(x, y){
loess1 <- loess( y ~ x , span=0.10) # 10% smoothing span
smoothed1<- predict(loess1) 
plot( x, y, type="p" )
lines(x, smoothed1, col="red")
}
p <- p %>% arrange(MOCA_Total)
linearity(p$MOCA_Total, p$framing)


linearity(p$MOCA_Total, p$MotorExamination_PartIIIScore)

linearity(p$MOCA_Total, p$HDRSTotScore)

# aim 2
linearity(p$MOCA_Total, p$avg_SBP)

linearity(p$MOCA_Total, p$avg_DBP)

linearity(p$MOCA_Total, p$Age)

linearity(p$MOCA_Total, p$BMI)

```

   
###Check Collinearity
```{r}
#continuous variables:  
  # continuous normal: age, height, weight, sbp, dbp
  # continous skew: depression score, motor score
checkcoll <- p

collinearity <- function(x, y){
loess1 <- loess( y ~ x , span=0.10) # 10% smoothing span
smoothed1<- predict(loess1) 
plot( x, y, type="l" )
lines(x, smoothed1, col="red")
}
```
#### Aim 1: Framingham Scores, PD/control, education, motor impairment score, Hamilton Depression Scale

```{r}
# framing with motor score
checkcoll$framing100 <- checkcoll$framing * 100 # rescale the framing score
checkcoll <- checkcoll %>% arrange(framing100)
collinearity(checkcoll$MotorExamination_PartIIIScore, checkcoll$framing100)
fit <- lm(MotorExamination_PartIIIScore ~ framing100 , data = checkcoll) #significant
summary(fit)

# framing with Hamilton Depression Rating Total Score
collinearity(checkcoll$framing100, checkcoll$HDRSTotScore)
fit <- lm(HDRSTotScore ~ framing100 , data = checkcoll) 
summary(fit)


# motor impairment score with Hamilton Depression Score
checkcoll <- checkcoll %>% arrange(MotorExamination_PartIIIScore)
collinearity( checkcoll$MotorExamination_PartIIIScore, checkcoll$HDRSTotScore)
fit <- lm(HDRSTotScore ~ MotorExamination_PartIIIScore , data = checkcoll) #significant?
summary(fit)
```


#### Aim 2: average SBP, average DBP, BMI, ever smoke, diabetes, hypertension medication, age, PD/control, gender, education, motor impairment score, Hamilton Depression Scale
```{r}
# 1.
#SBP with BMI
checkcoll <- checkcoll %>% arrange(BMI)
collinearity(checkcoll$BMI, checkcoll$avg_SBP) 
fit <- lm(avg_SBP ~ BMI , data = checkcoll) #significant
summary(fit)

#DBP with BMI
collinearity(checkcoll$BMI, checkcoll$avg_DBP) 
fit <- lm(avg_SBP ~ BMI , data = checkcoll) #significant
summary(fit)

#age with BMI
collinearity(checkcoll$BMI, checkcoll$Age) 
fit <- lm(Age ~ BMI , data = checkcoll) #significant
summary(fit)

# moter impairment score with BMI
collinearity(checkcoll$BMI, checkcoll$MotorExamination_PartIIIScore)
fit <- lm( MotorExamination_PartIIIScore~ BMI , data = checkcoll)
summary(fit)


#Hamilton Depression Score with BMI
collinearity(checkcoll$BMI, checkcoll$HDRSTotScore)
fit <- lm( HDRSTotScore ~ BMI , data = checkcoll)
summary(fit)

# 2.
## DBP with SBP
checkcoll <- checkcoll %>% arrange(avg_SBP)
collinearity(checkcoll$avg_SBP, checkcoll$avg_DBP)
fit <- lm(avg_DBP ~ avg_SBP, data = checkcoll) #significant
summary(fit)

# age with SBP
collinearity(checkcoll$avg_SBP, checkcoll$Age)
fit <- lm(Age ~ avg_SBP, data = checkcoll) #significant
summary(fit)

## moter impairment score with SBP
collinearity(checkcoll$avg_SBP, checkcoll$MotorExamination_PartIIIScore)
fit <- lm(MotorExamination_PartIIIScore ~ avg_SBP, data = checkcoll)
summary(fit)

#Hamilton Depression Score with SBP
collinearity(checkcoll$avg_SBP, checkcoll$HDRSTotScore)
fit <- lm( HDRSTotScore~ avg_SBP, data = checkcoll)
summary(fit)

# 3.
## age with DBP
checkcoll <- checkcoll %>% arrange(avg_DBP)
collinearity(checkcoll$avg_DBP, checkcoll$Age)
fit <- lm( Age~ avg_DBP, data = checkcoll) #significant
summary(fit)

# moter impairment score with DBP
collinearity(checkcoll$avg_DBP, checkcoll$MotorExamination_PartIIIScore)
fit <- lm(MotorExamination_PartIIIScore ~ avg_DBP, data = checkcoll) #significant
summary(fit)

#Hamilton Depression Score with DBP
collinearity(checkcoll$avg_DBP, checkcoll$HDRSTotScore)
fit <- lm(HDRSTotScore ~avg_DBP, data = checkcoll) 
summary(fit)

# 4.
checkcoll <- checkcoll %>% arrange(Age)
# moter impairment score with age
collinearity( checkcoll$Age, checkcoll$MotorExamination_PartIIIScore)
fit <- lm(HDRSTotScore ~ Age, data = checkcoll) 
summary(fit)

#Hamilton Depression Score with age
collinearity(checkcoll$Age, checkcoll$HDRSTotScore)


```

##Regression Analysis

###Aim 1: find out whether individuals with high Framingham score have greater cognitive impairment reflected by MoCA score in PD and control group.
predictor: Framingham Scores, PD/control, education, motor impairment score, Hamilton Depression Score



```{r}

fit5 <- lm(MOCA_Total ~ framing  + NeuroExamPrimaryDiagnos  + factor(edu) + MotorExamination_PartIIIScore  + HDRSTotScore , data = p) 
summary(fit5)


fit4 <- lm(MOCA_Total ~ framing  + NeuroExamPrimaryDiagnos  + factor(edu) + MotorExamination_PartIIIScore, data = p) 
summary(fit4)
lrtest(fit5, fit4)

fit3 <- lm(MOCA_Total ~ framing  + NeuroExamPrimaryDiagnos + factor(edu) , data = p) 
summary(fit3)
lrtest(fit5, fit3) 

fit2 <- lm(MOCA_Total ~ framing + NeuroExamPrimaryDiagnos , data = p)
summary(fit2)
lrtest(fit5, fit2) # likelihood ratio test

fit1 <- lm(MOCA_Total ~ framing, data = p)
summary(fit1)
lrtest(fit5, fit1)
#final model is fit5 

##### consider possible interaction with motor score
fit6 <- lm(MOCA_Total ~ framing *MotorExamination_PartIIIScore  + NeuroExamPrimaryDiagnos  + factor(edu)  + HDRSTotScore , data = p) 
summary(fit6)
lrtest(fit5, fit6) # no significance 

fit7 <- lm(MOCA_Total ~ framing  + NeuroExamPrimaryDiagnos *MotorExamination_PartIIIScore   + factor(edu)  + HDRSTotScore , data = p) 
summary(fit7)
lrtest(fit5, fit7) # no significance 

fit8 <- lm(MOCA_Total ~ framing  + NeuroExamPrimaryDiagnos  + factor(edu) * MotorExamination_PartIIIScore  + HDRSTotScore , data = p) 
summary(fit8)
lrtest(fit5, fit8) # no significance 

fit9 <- lm(MOCA_Total ~ framing  + NeuroExamPrimaryDiagnos  + factor(edu) + HDRSTotScore * MotorExamination_PartIIIScore , data = p) 
summary(fit9)
lrtest(fit5, fit9) #significant? interaction coefficient too small.

######### consider possible interaction with PD/Control
fit10<- lm(MOCA_Total ~  framing * NeuroExamPrimaryDiagnos + factor(edu) + MotorExamination_PartIIIScore + HDRSTotScore, data = p) 
summary(fit10)
lrtest(fit5, fit10) # no significance

fit11<- lm(MOCA_Total ~  framing  + factor(edu) * NeuroExamPrimaryDiagnos+ MotorExamination_PartIIIScore + HDRSTotScore, data = p) 
summary(fit11)
lrtest(fit5, fit11)#significant


fit12<- lm(MOCA_Total ~  framing  + factor(edu)+ MotorExamination_PartIIIScore + HDRSTotScore * NeuroExamPrimaryDiagnos, data = p) 
summary(fit12)
lrtest(fit5, fit12) # no significance

######### consider possible interaction with edu
fit13<- lm(MOCA_Total ~  framing  * factor(edu)+ MotorExamination_PartIIIScore + HDRSTotScore + NeuroExamPrimaryDiagnos, data = p) 
summary(fit13)
lrtest(fit5, fit13)# no significance

fit14<- lm(MOCA_Total ~  framing + NeuroExamPrimaryDiagnos  + MotorExamination_PartIIIScore  + HDRSTotScore * factor(edu) , data = p) 
summary(fit14)
lrtest(fit5, fit14) #no significance

######### consider possible interaction with Hamilton depression score
fit15<- lm(MOCA_Total ~  framing * HDRSTotScore + NeuroExamPrimaryDiagnos + factor(edu) + MotorExamination_PartIIIScore , data = p) 
summary(fit15)
lrtest(fit5, fit15)  #no significance

```

###MODEL DIAGNOSIS

```{r}
par(mfrow=c(2,2))
plot(fit5) #scale-location plot: violation of equal variance 
plot(fit9)
plot(fit11)
# test for equal variance (Score Test For Non-Constant Error Variance)
library(car)
ncvTest(fit9)
# pair-wise correlation
library(GGally)
ggpairs(X)  #the correlation between motor score and PD/Control is relatively high, 0.717

# Farrar – Glauber Test  - con
library(mctest)
X <- p[,c(3,6,8,9,18)]
X$NeuroExamPrimaryDiagnos <- ifelse(X$NeuroExamPrimaryDiagnos == "Control", 0, 1)
#X$edu <- factor(X$edu)
omcdiag(X,p$MOCA_Total)
imcdiag(X,p$MOCA_Total, method="VIF")
library(ppcor)
pcor(X, method = "pearson")$p.value <= 0.05

```
###Aim 2: identify the relationship between vascular risk factors (hypertension, obesity and smoking) and cognitive impairment. 
predictors: average SBP, average DBP, BMI, ever smoke, diabetes, hypertension medication, age, PD/control, gender, education, motor impairment score, Hamilton Depression Scale

```{r}


aim2fit1<- lm(MOCA_Total ~ avg_SBP + avg_DBP + BMI + factor(Smoke_Ever) + factor(diabetes) + factor(HTN_meds) + Age + NeuroExamPrimaryDiagnos  +  factor(Gender) + factor(edu) + HDRSTotScore + MotorExamination_PartIIIScore, data = p) #
summary(aim2fit1)

aim2fit2<- lm(MOCA_Total ~ avg_SBP + avg_DBP + factor(Smoke_Ever) + factor(diabetes) , data = p) #
summary(aim2fit2)

```


